### 2023-03-27 analysis/98-scrap.qmd
Created a file for abandoned words.
Initially containts work-in-progress from earlier un-committed iterations of the introduction, and I may draw from this as I continue work on the background section.

The ability to write code is of course fundamental to effective use of digital tools and media. However, as with many so-called "soft skills" that are crucial for academic professional development, additional competencies relating to the maintenance, management, and distribution of software, such as the ability to receive and implement feedback, set and stick with long-term goals, coordinate labour, document work practices, and collaborate with others, are underappreciated aspects of software development. This is particularly true pertaining to open source projects, which entail publishing source code on the web for free, maintaining transparent programming practices, and inviting community participation and feedback. Open source projects, which have been heralded as revolutionary, community-oriented, and anti-capitalist means of production, literally run the web (and, by extension, the modern global economy), but they are not the asocial, meritocratic systems that many proponents claim they are. In fact, open source projects incorporate complex organizational structures, governance strategies, and forms or social mediation to help delegate and vet contributions made by distributed participants. They rely on, rather than eschew, institutional support structures, in order to motivate work, keep volunteer  maintainers involved, and generally ensure that the project can be sustained over the long-term. Open source is more than just putting your code online; when done right, it is a social experience.

The adoption of open source development models among archaeologists is generally informed by the broader open science movement, which is motivated by a desire to facilitate novel research opportunities, to make participation in scienctific research more equitable, to reclaim science as a public good, and to enhance the means of validating findings. Open science advocates have put a lot of effort into developing information infrastructures, or systems designed to facilitate the transmission of data on a global scale, but, at least among archaeologists, their transformative value has not yet been realized. For instance, numerous studies have highlighted how archaeologists are not reusing data published on open data platforms, and instead opt to communicate directly with those who created the dataset to obtain a relationship that more closely resembles a more traditional form of collaboration. In other words, these pltforms do little to alter the professional norms and expectations that frame archaeological research. They provide the capacity and the potential to enable change, but do not in themselves change the ways in which archaeologists create and value research outcomes. This must come from community buy-in, and the extension and establishing of novel forms of collaborative commitments that reflect the ideal state of affairs. In other words, it is a social problem, not a technical one.

---

Archaeologists have been developing specialized software since the 1960s, when computers were first introduced in the social science (cf. Cowgill). The popular adoption of personal computers has contributed to the growth of digital archaeology as a distinct field of practice. Although computer programming is not generally taught in archaeological curricula, archaeologists are generally eager to apply novel technologies to address their research questions.

Software development entails designing, programming, distributing, and maintaining software, and is a key aspect of contemporary digital archaeology.

* There is open archaeology
The open science movement has emerged as a series of professional practices that are aimed at making research more accessible and transparent.

The term 'open' is somewhat nebulous, but a generaly agreed upon definition defines open as "A piece of content or data is open if anyone is free to use, reuse, and redistribute it – subject only, at most, to the requirement to attribute and/or share-alike."

This definition of open does not state what open is for, or what motivates the open science movement. However, a brief reading of open science manifestos leads us to believe that open is about a few things: facilitating novel research opportunities, making participation in scienctific research more equitable, reclaiming science as a public good, and enhancing validation of findings.

A lot of work has gone in to developing information infrastructures, or systems designed to facilitate the transmission of data on a global scale. Open data repositories, open peer-review platforms, and online social spaces have been set up to foster participation in new forms of scholarly communication.

- By making research materials and processes more accessible and transparent, it is believed that scientific knowledge production can be ehanhced.
- In part, this is through the increased capacity to use these materials, i.e. it is an efficiency concern. Simply having these materials available is a fundamental requirement to facilitate their use. It is about reducing barriers, in terms of information transmission and legal accomodations.
- As such, a lot of attention goes into developing technical protocols and standards to facilitate data sharing and reuse
- But, as some studies have noted, these infrastructures often fail to deliver on these promises, and casual observation indicates how people value connections that these platforms fail to support
- These issues centre on the inability to account for social and professional expectations
- Either the professional culture has not cought up with the technical infrastructures (which impose novel sociotechnical arrangeemnts), or the infrastructures serve needs that archaeologists do not feel they actually need
- What is the point of establishing these infrastructures, and in making research materials available, if no one is going to use them, or use them in a way that actually serves their intended purpose?

---

For instance, the Linux kernel is generally dependent on a coterie of senior developers, who are financially supported by the Linux Foundation, who delegate and vet contributions made by unpaid volunteers. Simularly, the Firefox web browser depends on funding from Alphabet, the corporate entity which dominates the space in which they operate, to stay afloat.

Open source software development models are crucial for maintaining key software that runs the internet (and, by extension, the modern global economy), and have been generally posed as revolutionary, community-oriented, anti-capitalist means of production. But even the most prominent open source projects contain bottlenecks that disrupt these idealistic imaginaries of heterarchical organizational structures, meritocratic decision-making, and potential for multivocal and community-led growth. For instance, the Linux kernel is generally dependent on a coterie of senior developers who delegate and vet contributions (cf. Cyberchiefs), the Firefox web browser is dependent on funding from the corporate entity which dominates the space in which they operate, and the massive distributed web of libraries and extensions that make up various software development ecosystems (such as JavaScript, Python and R) struggles to keep maintainers interested in continuing to provide support on a volunteer basis (as illustrated by the left-pad incident). Open source is not a panacea for institutional inertia and inequity, and actually depends on social relations and governance strategies to remain sustainable (Cyber chiefs, two bits, Matt's dissertation).

What does this mean for the adoption of open source ideas and practices among archaeologists? Archaeologists work in a very different institutional space than these very prominent projects, and are influenced by different profesional pressures, such as the warrant to publish, inadequate access to long-term research funding, the obligation to train students, the need to work as part of university systems with intricate institutional budgeting practices (which pit likeminded teams of researchers against each other), and the need to move on from projects after producing stable research outputs. At first glance, many of these aspects of being an archaeologist are in opposition with open source strategies. These tensions lead us to wonder how archaeologists actually implement open source. Beyond simply making their code available on the web, do they also implement the social strategies that make open source productive? Does archaeological open source actually help achieve greater transparency, sustainability, and community participation? And if not, what does it actually achieve?