### 2023-03-28 analysis/05-history_of_open_archaeology.qmd

Starting with a relatively simple plot, this is a chart of the cumulative growth of open archaeology over the years, measured by code contributions committed and pushed to GitHub repositories.

As we can see, archaeologists have been using git from even before GitHub was launched in 2008. But they really began to get into it around 2014--2015, when we see an uptick in the rate of growth that has continued ever since.

Around this time we also see that GitHub starts being used to host documents and scripts. This may represent a recognition of GitHub's ability to track things other than code, and a willingness to experiment with version control systems as a medium for disseminating work, especially work or manifestations of undervalued skills that may not necessarily be easily represented using traditional publication venues, in an open and somewhat nerdy way.

{r plot-gh-cum}

---

- Growth of open archaeology by number of commits
- Growth of open archaeology by number of repos
- Compare changing proportions of each category on a year by year basis
- Identify periods associated with different rates of growth, and relate these with different general attitudes in the history of digital archaeology