### 2023-03-28 analysis/06-project_activity.qmd

In this section we assess projects' pulse, or the amount of activity pertaining to a repository. In the following figure, each vertical line represents the lifespan of a repository, and the points on the line are individual commits. The graph is sorted by the date of the most recent commit for each repository, so the ones on the top became inactive earlier and the ones on the bottom represent ongoing work.

{r plot-lifespan}

The span of time between the first and latest commits is visualized more clearly in summary here. Notably, many projects are active for only a few short period of time: about 15% less than a month, 25% less than three months, and 40% less than a year. Projects with a short span of activity tend to have very high commit frequency, whereas there is much more variety in long-term projects. Among projects with a greater lifespan, those with more total commits have much more frequent activity than those with fewer total commits, which represent the majority of long-lived projects. The overall negative trend (seen in the top-right graph) can be explained in terms of natural processes, i.e. the only way to fit a high number of commits within the same span as a project with a low number of commits is to commit more frequently. But bearing in mind both axes use a logarithmic scale, the upper left portion of the graph does appear to suggest that short-lived projects are characterised by a spree of rapid committing; while longer-lived projects have a more diverse range of workrates. This latter pattern in turn appears to be explained by the number of contributors: projects that sustain a high commit rate over a long period of time tend to rely on a large number of contributors, as per the bottom-right graph.

{r lifespan-stats}

{r plot-lifespan-hist}

{r plot-lifespan-vs-rate}

Perhaps unfortunately then, we also find that nearly 60% of projects have only a single contributor, and projects with more than a dozen are very rare indeed \[\<3%\]. A "contribution", in this case, is a commit, issue, comment applied by a user to a repository.

{r plot-gh_contributors-histogram}

Furthermore, when we break down the contributions made by all contributors in multi-contributor projects, we also see that work tends to be distributed very unevenly. Of multi-contributor repositories, 92% have a third of contributions made by a single person; 65% have half of all contributions made by a single person; and 36% have three quarters of all contributions made a single person.

{r plot-contrib-dist}

Perhaps as a result of this, we also find that only a minority of projects make much use of collaborative features of GitHub such as issues and pull requests (left) and comments (right). Only 39% of repositories make use of these features. Of these repositories with any issues and pull requests, 28% have only 1 and 82% have less than 10. However, of all repositories that have any comments posted in reply to issues and pull requests, there is actually a significant degree of activity. This seems to suggest that people are willing to respond to issues raised by others, but may not necessarily want to take initiative, for a variety of reasons that may be more effectively investigated through qualitative analysis of the conversations in the issue threads.

{r plot-issues-hist}

{r plot-comments-hist}