### 2023-03-29 analysis/02-background.qmd
Initial and probably overly ranty draft of the background section.

* What is open source?

Open source is a software development model that prioritizes transparent work processes. Initially driven by the idea that computer users should be free to understand and manipulate the software that they install on their computers (e.g., "free software", as initially conceived by the Free Software Foundation), open source has become a means of collaborative software development (Kelty 2008: chapter 3, especially starting at page 99). By putting one's code on the web without restriction on how it may be used or manipulated, this encourages creativity to flourish as people contribute to help improve the code base. Software thus emerges from the coordinated labour of worldwide volunteers, who shape the product according to the collective vision. An open code base may also be used to support alternative projects whose missions diverge from the original plan, and an entire project may be "forked", or taken in a new direction if contributors are dissatisfied with how core developers run things.

Open source has traditionally been referred to as being based on meritocratic principles. A good test of whether a contribution should be included in a published software release is whether it is functional. Moreover, with more eyes looking over a code base, it is easier to identify flaws with a contribution and flag potential bugs or security issues. This is all done in the spirit of producing functional code, and in ideal circumstances faulty contributions will be corrected before inclusion. Personal ego is minimized in favour of co-creating stable and functional outcomes. (Coleman 2012)

However, this is not the same as saying that open source is completely anarchic or based on the "wisdom of crowds". In fact, successful open source projects incorporate complex organizational structures, governance strategies, and forms or social mediation to help delegate and vet contributions made by distributed participants. They rely on, rather than eschew, institutional support structures, in order to motivate work, keep volunteer maintainers involved, and generally ensure that the project can be sustained over the long-term. Open source is more than just putting your code online; to be successful, it requires participation in a social experience. (Kelty 2008; O'Neill 2009; Ratto dissertation and derived works)

In other words, as with many so-called "soft skills" that are crucial for academic professional development, additional competencies relating to the maintenance, management, and distribution of software, such as the ability to receive and implement feedback, set and stick with long-term goals, coordinate labour, document work practices, and collaborate with others, are grossly underappreciated factors that contribute to an open source project's success.

We therefore consider open source to be a means of collaboration more than a means of transmitting information. It involves developing software as part of a group, developing consensus, and working with common purpose. Crucially, it also involves having a welcoming attitude, a sense of humility, and an understanding that one's work may be appropriated and used in unanticipated ways.

* How does open source relate to open science?

The open science movement comprises a series of practices and principles intended to make research more accessible, transparent, and efficient. Although the concept of "open" is somewhat nebulous, in terms of its abstract definition and with regards to what real-world applications count as being open, one commonly-cited definition describes content that "can be freely used, modified, and shared by anyone for any purpose" (https://opendefinition.org/). This definition does not state what open is for, how to be open, or any sort of social or discursive framing behind the open movement. However, most open science advocates claim that they are motivated by a desire to facilitate novel research opportunities, make participation in scienctific research more equitable, reclaim science as a public good, and enhance how findings are validated and legitimized.

The idea that scientistis should generally contribute to a public domain of knowledge without profit motive has led to open science being heralded as revolutionary, community-oriented, and anti-capitalist means of production. However, while open science does have the _potential_ to effect radical change, this is not a given. The social and institutional contexts in which we do science is firmly embedded within capitalist and neoliberal power structures that reward individualistic competition and do little to actually encourage equitable and accessible research practices, and as such, make it difficult to fully embrace open science ideals. Moreover, the open science movement, which is dominated by STEM disciplines, prioritizes a grossly simplified and asocial notion of what science is and entails. Namely, it considers science as the accumulation and assembly of a species-level understanding of the world, which is not held by any one individual but is stored in seemingly value-neutral and disembodied media, facts and observations. This is manifested by digital telecommunications systems that host files, document processes, facilitate co-working opportunities, and perform automated processes. However, these systems have become so emblematic of open science that the use of these tools and resources designed to _support_ open science is often mistaken for _actually doing_ open science.

Open science is typically compared with the open source movement in that they both involve a distributed, digitally-mediated and worldwide labour force, who somehow derive rough consensus directed towards assets held in the public domain. But they differ in terms of the contexts in which they operate, the stakeholders involved, and the kinds of outcomes they produce. Whereas open source emerged from concern of consumer rights and then developed as a means of maintaining resilient and collectively motivated projects, open science comes out of a desire to make research practices more transparent and accessible. Open source is performed by professional and hobbyist software developers alike, and participants contribute in a wide variety of ways (inlcuding: programming, writing documentation, translating software and documentation, bug reporting, and financial support), but, in open science, scientists are usually the only participants actively involved in creating and maintaining contributions. Moreover, whereas open source projects often attract participants with varied stakes in the software and use cases in mind, open science projects are typically bounded by small communities of specialists with very particular needs, and rarely anticipate use of their work beyond the limited scope of their primary targeted audience. Additionally, open science is bounded by the professional contexts in which science operates, and as such, produces outputs that can be easily credited to specific sets of individuals for reasons of tenure and promotion. Open science projects whose contributions are supported by research funding also face sustainability concerns, as participants lose motivation to contribute once funding runs out. Once a project is completed and credit has been allocated, it is common for scientists to mark their projects as completed and move onto new endeavours. Open source projects, on the other hand, are motivated by a more practical need for the software to function properly, and contributors may remain actively or sporadically involved to satisfy users' needs.

* Situating open archaeology in open science

The "archaeological toolbox" is a common metaphor that describes the set of physical and informational tools and processes that archaeologists rely on to achieve their goals. In applying either physical or digital methods, archaeology has a long tradition of adopting common or "off the shelf" tools to serve their own unique purposes, which are not typically unanticipated by their original designers. In other words, the apparatus of archaeology is an open system that relies on products and skillsets from outside the discipline, but which exhibit enough flexibility to be used in novel ways. For example, archaeologists are content to outsource the metallurgical skills needed to forge their trowels, and delegate the manufacture and acquisition of key electronic tools, like laptop computers and cameras, to the global supply chains and labour networks that produce consumer-grade products for purchase. However, an increasingly large number of archaeologists are taking it upon themselves to program their own software, rather than rely on generic and commercially-produced information processing systems. In fact, archaeologists are prolific computer programmers and have been coding a variety of functions and libraries that serve uniquely archaeological problems and purposes. This software is typically made to support tasks that archaeologists find uniquely meaningful, and which are not supported by mainstream technology companies. This differs from the use of purpose-specific scripts written to address specific research questions. Such work often falls under the purview of open science, since making one's scripts openly available is said to enhance the potential for reproducibility (Marwick and Schmidt).

The adoption of open source development models among archaeologists is generally informed by the broader open science movement, which is motivated by a desire to facilitate novel research opportunities, to make participation in scienctific research more equitable, to reclaim science as a public good, and to enhance the means of validating findings. Open science advocates have put a lot of effort into developing information infrastructures, or systems designed to facilitate the transmission of data on a global scale, but, at least among archaeologists, their transformative value has not yet been realized. For instance, numerous studies have highlighted how archaeologists are not reusing data published on open data platforms, and instead opt to communicate directly with those who created the dataset to obtain a relationship that more closely resembles a more traditional form of collaboration (Huggett 2018; Atici 2013; Faniel 2013; Sobotkova 2018; Opitz 2021). In other words, these pltforms do little to alter the professional norms and expectations that frame archaeological research, and which seem to be fundamental factors that influence the nature of archaeological research. Open data platforms may provide the capacity and the potential to enable change, but do not in themselves change the ways in which archaeologists create and value research outcomes. This must come from community buy-in, and organic extension and establishing of novel forms of collaborative commitments that reflect the ideal state of affairs. In other words, it is a social problem, not a technical one.

Similarly, archaeological software development is framed by the circumstances under which research that supports and/or motivates the effort occurs. This means that software development processes are shaped by the pressures and circumstances of being an archaeologist, such as the need to publish and to generate stable research outputs, archaeologists' tendency to move on from projects after they are 'completed' (i.e. when money runs out and people lose interest), the need to be recognized as a citable author for purposes of hiring and promotion, and the general lack of formal digital literacy among archaeologists (which either produces a digital archaeology subdiscipline, a kind of ad-hoc cobbled together system (made by the most digitally literate person in the room), or cross-disciplinary connections with non-archaeologists (spouses, childhood friends, maybe a computer scientists, rarely someone in private industry)). (Morgan 2015; Morgan and Eve 2012; Batist et al 2021; Caraher 2022; Graham 2020; maybe also Thorpe 2012 and Zorzin 2015 on economic models that frame (non-digital) archaeological research)

A lot of work has gone in to developing information infrastructures, or systems designed to facilitate the transmission of open materials on a global scale, including open data repositories, open peer-review platforms, and online social spaces meant to foster participation in new forms of scholarly communication. However, the general failure of open science to deliver on its lofty promises has led archaeologists to reflect on what openness really means and entails. In the past, open archaeology was very much wrapped up in discourse on the best tools to use, optimal data processing pipelines, and tying into global, web-based infrastructures, protocols and standards (e.g., establishing proper and secure web servers, open api practices, using orcid, proper licencing, using markdown, git, R package development practices, etc) (cf. Kansa and Kansa 2014; Kintigh 2014; Roosevelt 2015; others), and this continues to be a prominent part of discourse on the development of open infrastructures (cf. Kansa on open identifiers; others). However, more recent reflections have attempted to re-frame research practices, particularly those corresponding with "data work", as situated experiences and associated with a continuum of decisions made within pragmatic contexts of action. Archaeological research is thus framed as inherently collaborative, and the materials and media that archaeologists rely on to do their research are framed as discursive entities, or communicatve media that bridge a gap of experience or understanding. Bridging this gap entails an act of reconciliation between different pragmatic experiences, and the technical infrastructures we have set up to handle these acts of communication do not adequately account for the range of factors that must be considered to successfully pass this reconciliation.