### 2023-03-30 analysis/02-background.qmd
Re-wrote the third section with a more positive tone

## What is open source?

Open source is a software development model that prioritizes transparent work processes. Initially driven by the idea that computer users should be free to understand and manipulate the software that they install on their computers (e.g., "free software", as initially conceived by the Free Software Foundation), open source has become a means of collaborative software development (Kelty 2008: chapter 3, especially starting at page 99). By putting one's code on the web without restriction on how it may be used or manipulated, this encourages creativity to flourish as people contribute to help improve the code base. Software thus emerges from the coordinated labour of worldwide volunteers, who shape the product according to the collective vision. An open code base may also be used to support alternative projects whose missions diverge from the original plan, and an entire project may be "forked", or taken in a new direction if contributors are dissatisfied with how core developers run things.

Open source has traditionally been referred to as being based on meritocratic principles. A good test of whether a contribution should be included in a published software release is whether it is functional. Moreover, with more eyes looking over a code base, it is easier to identify flaws with a contribution and flag potential bugs or security issues. This is all done in the spirit of producing functional code, and in ideal circumstances faulty contributions will be corrected before inclusion. Personal ego is minimized in favour of co-creating stable and functional outcomes. (Coleman 2012)

However, this is not the same as saying that open source is completely anarchic or based on the "wisdom of crowds". In fact, successful open source projects incorporate complex organizational structures, governance strategies, and forms or social mediation to help delegate and vet contributions made by distributed participants. They rely on, rather than eschew, institutional support structures, in order to motivate work, keep volunteer maintainers involved, and generally ensure that the project can be sustained over the long-term. Open source is more than just putting your code online; to be successful, it requires participation in a social experience. (Kelty 2008; O'Neill 2009; Ratto dissertation and derived works)

In other words, as with many so-called "soft skills" that are crucial for academic professional development, additional competencies relating to the maintenance, management, and distribution of software, such as the ability to receive and implement feedback, set and stick with long-term goals, coordinate labour, document work practices, and collaborate with others, are grossly underappreciated factors that contribute to an open source project's success.

We therefore consider open source to be a means of collaboration more than a means of transmitting information. It involves developing software as part of a group, developing consensus, and working with common purpose. Crucially, it also involves having a welcoming attitude, a sense of humility, and an understanding that one's work may be appropriated and used in unanticipated ways.

## How does open source relate to open science?

The open science movement comprises a series of practices and principles intended to make research more accessible, transparent, and efficient. Although the concept of "open" is somewhat nebulous, in terms of its abstract definition and with regards to what real-world applications count as being open, one commonly-cited definition describes content that "can be freely used, modified, and shared by anyone for any purpose" (https://opendefinition.org/). This definition does not state what open is for, how to be open, or any sort of social or discursive framing behind the open movement. However, most open science advocates claim that they are motivated by a desire to facilitate novel research opportunities, make participation in scienctific research more equitable, reclaim science as a public good, and enhance how findings are validated and legitimized.

The idea that scientistis should generally contribute to a public domain of knowledge without profit motive has led to open science being heralded as revolutionary, community-oriented, and anti-capitalist means of production. However, while open science does have the _potential_ to effect radical change, this is not a given. The social and institutional contexts in which we do science is firmly embedded within capitalist and neoliberal power structures that reward individualistic competition and do little to actually encourage equitable and accessible research practices, and as such, make it difficult to fully embrace open science ideals. Moreover, the open science movement, which is dominated by STEM disciplines, prioritizes a grossly simplified and asocial notion of what science is and entails. Namely, it considers science as the accumulation and assembly of a species-level understanding of the world, which is not held by any one individual but is stored in seemingly value-neutral and disembodied media, facts and observations. This is manifested by digital telecommunications systems that host files, document processes, facilitate co-working opportunities, and perform automated processes. However, these systems have become so emblematic of open science that the use of these tools and resources designed to _support_ open science is often mistaken for _actually doing_ open science.

Open science is typically compared with the open source movement in that they both involve a distributed, digitally-mediated and worldwide labour force, who somehow derive rough consensus directed towards assets held in the public domain. But they differ in terms of the contexts in which they operate, the stakeholders involved, and the kinds of outcomes they produce. Whereas open source emerged from concern of consumer rights and then developed as a means of maintaining resilient and collectively motivated projects, open science comes out of a desire to make research practices more transparent and accessible. Open source is performed by professional and hobbyist software developers alike, and participants contribute in a wide variety of ways (inlcuding: programming, writing documentation, translating software and documentation, bug reporting, and financial support), but, in open science, scientists are usually the only participants actively involved in creating and maintaining contributions. Moreover, whereas open source projects often attract participants with varied stakes in the software and use cases in mind, open science projects are typically bounded by small communities of specialists with very particular needs, and rarely anticipate use of their work beyond the limited scope of their primary targeted audience. Additionally, open science is bounded by the professional contexts in which science operates, and as such, produces outputs that can be easily credited to specific sets of individuals for reasons of tenure and promotion. Open science projects whose contributions are supported by research funding also face sustainability concerns, as participants lose motivation to contribute once funding runs out. Once a project is completed and credit has been allocated, it is common for scientists to mark their projects as completed and move onto new endeavours. Open source projects, on the other hand, are motivated by a more practical need for the software to function properly, and contributors may remain actively or sporadically involved to satisfy users' needs.

## Open archaeology at a crossroads

The adoption of open source development models among archaeologists is generally informed by the broader open science movement, which is motivated by a desire to facilitate novel research opportunities, to make participation in scienctific research more equitable, to reclaim science as a public good, and to enhance the means of validating findings. Open science advocates, including archaeologists, have put a lot of effort into developing information infrastructures, or systems designed to facilitate the transmission of data on a global scale. And as with open science in general, open archaeology has been wrapped up in discourse on the best tools to use, optimal data processing pipelines, and tying into global, web-based infrastructures, protocols and standards (cf. Kansa and Kansa 2014; Kintigh 2014; Roosevelt 2015). This continues to be a prominent part of discourse on the development of open infrastructures (cf. Kansa 2022 on unique  identifiers; something about aDNA pipelines; Hacıgüzeller 2021 on CIDOC-CRM (though she is actually very critical); Huvila and his team on formal paradata), but more recent reflections on the use of these infrastructures recognize underlying and systemtic commitments and assumptions that inform how these systems are used and valued by various stakeholders. For instance, the pressures and circumstances of being an archaeologist, such as the need to publish stable research outputs, and archaeologists' tendency to move on from projects after they are completed, contribute towards shaping data sharing platforms as delegated archival services, whereupon the obligation to care for data over the long-term is passed on to others. In response to criticisms that highlight how these systems add little practical downstream value (Huggett 2018; Sobotkova 2018), open archaeology advocates have begun to prioritize the role that information infrastructures may play in contexts of data reuse, namely by implementing the FAIR principles in order to ensure that data are findable, accessible, interoperable, and reusable (Kansa 2023 in that recent special issue in Advances in Archaeological Practice; one or two additional sources). This example is matched by similar efforts to raise the value proposition of using open science tools and systems, such as through an emphasis on enhancing R package development and documentation skills so that code may be run and evaluated in alternative computing environments (cite Marwick and/or Roe), and [another example].

Additionally, archaeologists are becoming more comfortable showcasing and actively discussing undervalued research ouputs such as software and datasets, which are reviewed and published by venues like Peer Community In, the Journal of Archaeological Data, and the Journal of Open Source Software. Professional blogs, social media and podcasts are also gaining traction as proper venues for disseminating one's work. This is consistent with archaeology's historical willingness to experiment with new methods and media and its casual attitude towards scholarly communication. Archaeologists' embrace of methodological plurality and their positions as interdisciplinary scholars scattered across institutional boundaries has prepared them for this kind of pragmatic, improvisatory approach to scholarly communication via the web. There is enormous potential for archaeologists to reinvent novel forms of scholarly communication and extend a new set of collaborative commitments that go beyond those typically recognized by scholarly institutions, and which correspond with open science principles. But this depends on archaeologists' willingness to continue experimenting and on the community's willingness to buy into novel value regimes.