### 2023-09-20 analysis/_07-conclusion.qmd
Merge outline into text

# Conclusion

<!--

There is an emerging community of practice around open source in archaeology.

* It is limited and fragmented
  * A small set of subfields are represented
  * Probably most people still haven't heard it?
* Collaboration is limited
  * Most work remains solo & short-lived
  * Perhaps because open source norms conflict with academic norms?
* The members of this community face structural challenges in academia
-->

Archaeology is a "magpie" discipline, using tools and methods that originate from a variety of other disciplines and industries to better understand the past. 
Moreover, archaeologists are accustomed to working on shoestring budgets, having to adapt their tools to suit specific, non-generalizable situations, and applying a DIY work ethic to create and co-creating with others to get the job done (Batist et al 2021; Morgan and Eve 2012; Caraher 2019). 
It is therefore unsprising that archaeologists are prolific open source software developers.

In this study we have sought to document an emerging community of open source software engineers in archaeology and to assess the extent to which collaboration within this community resembles conventional 'open source' forms. 
We have operationalised 'open-source collaborative practices' as the use of certain features of git and GitHub visible to us in data from the GitHub API. 
This excludes a wide range of other potential practices, most obviously the use of other software (though as we demonstrate in section X.X, the git/GitHub combination is overwhelmingly the most popular), but also collaboration through offline or private channels, or forms of collaboration we don't know about. 
We have not directly interviewed those doing the collaborating, though our conclusions do drawn heavily from our experience as members of that community ourselves. 
Our earliest data is from XXX and our study can say little about collaborative software development in archaeology before this point, though we know there was a significant amount of it (@cite).

These caveats notwithstanding, we have documented that open source software development in archaeology has seen a rapid and sustained rise beginning around 2014. 
This is marked by a variety of applications and use cases, including the use of git and GitHub to track and host content other than code. 
Moreover, archaeologists are very involved in broader scripting ecosystems, as is evident through the predominant creation of R packages and Python libraries designed to process the rich variety of archaeological information. 
At the same time, archaeologists also create standalone software for more intensive tasks that require greater access to system resources or that warrant more complex user interfaces than what R and Python IDEs are capable of providing.

Looking at the distribution of tags, we note that, among items listed as software, repositories tend to be focused on 
various means of identifying distribution patterns (spatial, temporal, statistical), 
calibrating data obtained from various instrumental methods (XRF, luminescence dating), 
supporting specialized finds analysis (zooarchaeology, palaeobotany, archaeogenetics), 
and supporting the collection and processing of archaeological materials.
These foci signify gaps in the archaeological toolbox that archaeologists recognized, and have attempted to fill, on their own terms.
In future studies, it would be interesting to examine how the purposeful expansion of open source tools corresponds with broader methodological trends apparent in publishing patterns or through other expressions of archaeological interest (e.g. social media posts, conference sessions).

While there is an emerging community of practice around open source in archaeology, we observed that collaboration remains limited. 
Most work is performed individually and is short-lived. 
The vast majority of repositories have 1--3 contributors, with only a few distinguished by an active and diverse developer base. 
Our analysis also shows an uneven use of git and GitHub's extended features, beyond their basic usage as a version control system and repository host. 
Generally speaking, we believe that this is because people do not want to step on other people's shoes by raising issues or intruding on other people's projects. 
This may relate to the fact that most developers on this list are academics who hold different values relative to the designers of open source development environments, regarding how collaboration should occur, for example, when dealing with how projects and ideas are 'owned' by individuals or communities, and how work should be iteratively improved upon.

However, we seem to have entered a period of critical reflection on the practical ramifications of infrastructures developed to support open science. 
For instance, Huggett (2022) identified social frictions as a necessary aspect of productive research, which open science infrastructures effectively seek to eliminate. 
Numerous other studies by Batist (2023), Huvila (2022), Haciguzeller et al. (2021), Opitz et al. (2021) Sobotkova (2018), Wylie (2017) Atici et al. (2013), Faniel et al. (2013) (among others) highlighted the need to account for the work involved in preparing and making use of materials shared openly on the web, and which are effectively muted by infrastructures developed in pursuit of open science.

Our network analysis similarly draws attention to the real-world collaborative ties that underpin archaeological open source software development. We identified three clusters of close associations between projects: 
the first is centred around a highly specialized supdiscipline exhibiting a very enthusiastic attitude (**ZB NOTE: or maybe even pressure?**) toward working in the open; 
the second is concerned with developing solutions to a very common and discipline-wide problem, specifically relating to data collection and management in fieldwork; 
and the third deals with several topic of concern, but represents an emerging community of practice. 
This third cluster comprises efforts performed by members of a close-knit community, many of whom already collaborate at local research centres or participate in relatively private interest groups. 
This indicates that open source is firmly embedded within existing social and institutional support and power structures that permeate academic life, both online and offline.