### 2024-03-21 analysis/_99-appendices.qmd
Add appendices

# Data availability {.appendix}

The results presented here are based on the directory of open archaeology projects maintained at <https://open-archaeo.info> [@batist2023].
The specific version we used is available from Zenodo at <https://doi.org/10.5281/zenodo.10625236> as well as in the compendium associated with this paper [@roe2024, https://zenodo.org/doi/10.5281/zenodo.8393043], in CSV format under `analysis/raw_data`.

Further data on activity in the git repositories associated with these projects was obtained via the GitHub API.
This data is available from Zenodo in the compendium associated with this paper [@roe2024, https://zenodo.org/doi/10.5281/zenodo.8393043], in RData and CSV format under `analysis/derived_data`.

# Acknowledgements {.appendix}

The impetus for this work was a session at the Computer Applications and Quantitative Methods in Archaeology (CAA) virtual annual meeting in Cyprus in June 2021, 'Tools for the Revolution: developing packages for scientific programming in archaeology'.
We are grateful to the conference organizers as well as to the (co-)convenors of the session, Martin Hinz and Clemens Schmid, for facilitating it.
Two anonymous reviewers offered feedback on our initial conference abstract.
The session was organized by the CAA special interest group on 'scientific scripting languages in archaeology' (CAA-SSLA, https://sslarch.github.io/), and we must thank its members—especially Sophie Schmidt, Clemens Schmid, Martin Hinz and Ben Marwick—for the encouragement in developing the work further.

Our original manuscript benefited greatly from the insights of the anonymous reviewer for *Internet Archaeology*.
We are also grateful to *IA* editor Judith Winters for managing the process and facilitating the interactive figures.

JR was supported by the Swiss National Science Foundation (SNSF Project #198152) for the duration of the production of this work.

<!-- TODO: Zack: anything to add? -->